\section*{Acknowledgements}
% \markboth{Acknowledgements}{Acknowledgements}

\epigraph{``Understanding is love's other name.''}{Thich Nhat Hanh}
\vspace{-20pt}
\epigraph{``We're philosophers. We think, therefore we am.''}{\textit{Small Gods}, Terry Pratchett}
\vspace{-20pt}
\epigraph{Films live or die on their casting.}{\href{https://youtu.be/W5bB7-jg_4M?list=PLomTW5B3fAMXJsqE_HDiMpdOdcfAk_Lqc&t=146}{Peter Jackson}}

I try as often as I can to remember that we live on a miraculous, miniscule blue dot floating in a vast expanse of emptiness, with scant resources and an incredible story.
%
It is the joy of my life to be part of the humanity sector of team earth.

\vspace{10pt}
\hrule

\begin{center}
Like most of the humanity sector, I am nothing without my mentors.
\end{center}


My first acknowledgements are to my tireless advisor, Jesse Thaler.
%
Thank you for taking me under your vast wing.
%
I have been very lucky to learn from your mastery over every aspect of the scientific process.
%
I have never seen someone do and change so much in the world in such a short time.



My second acknowledgments are to my parents, without whom I would surely not be here.
%
My mother taught me, through her shining example, the values of kindness, openness, acceptance, and sharp awareness and fighting spirit.
%
My father taught me by lack of example the importance of \glslink{jet-grooming}{grooming}, discussed further in \Chap{grooming}.
%
He also gave me the gifts of critical thinking, his jeans, and uncountable priceless lessons, experiences, and memories.



Tengiz Bibilashvili, my academic advisor during my time as an undergraduate at the University of California, Santa Barbara (UCSB), is one of the sharpest, kindest, brightest, and most generous human beings I have ever met.
%
I am grateful every day that I was lucky enough to experience his teaching.



Nathaniel Craig, my research advisor at UCSB, is a powerhouse of intellect, creativity, and pedagogy;
%
I benefited deeply from his dedication to mentorship and patience, and would have been lost on the tip of the iceberg of quantum field theory without his guidance (though, as he would always say with a smile, he was ``just doing his job'').



There is not enough I can say about the mentorship of Seth Koren.
%
Seth is incredibly empathetic and creative, and a brilliant teacher.
%
Seth is, to me, a quintessential human being:
%
a philosopher, clever and kind, who accepts nothing but the truth.



Finally, Wouter Waalewijn, who I met in the fourth year of my PhD, has been an incredibly patient mentor and a valuable resource for me in learning the ins and outs of quantum chromodynamics.
%
His presence is felt in many subtle ways throughout this thesis.
%
I can see his gentle, knowing smile even now.



I would also like to thank Iain Stewart and Christoph Paus.
%
They are members of my thesis committee, and I am grateful for their time and effort on my behalf, but I am most grateful for the environment they created at MIT through their presence.
%
They always had a warm comment to make or a laugh to share.
%
I knew I could always go to Iain with any question or concern and receive a listening ear and some sound advice.
%
I knew that I could always share a joke and a lighthearted chuckle with Christoph.


On the topic of environment, I would finally like to thank Daniel Harlow, who always had time to give a smile, kind words, and guidance in physics or beyond.

I am also extremely grateful for Paul Zeizel, who, along with his sharp wit and skill in listening, has the incredible talent of always saying the right thing.

\vspace{12pt}
\hrule
\vspace{12pt}


My sisters, Christina Nazmi-Williams and Mehrnoush Minor (\textit{in alphabetical order}), are my role models and favorite people.
%
It is thanks to their hard work and sacrifices on my behalf that I have been able to live such a clear and irresponsible life.
%
I am so lucky to be in their family.
%
Thomas Minor, my brother-in-law, is another favorite.
%
I aspire to someday be like Tom:
%
brilliant and humble, quietly aware and listening more than speaking, kind and gentle with a troublemaking streak.
%
My nephews and niece -- Ethan and Mason Williams, and Alex and Amelia Minor -- are bundles of joy and diamonds of my life.



My godparents Pieter and Martha van der Mersch, and my godsiblings Anna and Stephan, have provided my life with incredible warmth, amazing memories, and more empathy and useful guidance than I could have ever hoped for.



I also have the best siblings I could ever hope not to be related to.
%
Noah Mihelic (with credit and gratitude to my adopted family: Adrienne, Mark, and Josh Mihelic), who I knew before I had memories, has given me clarity, critical reasoning, incredible companionship, and many magical adventures.
%
David Rower (with credit and gratitude to my adopted family: Rachel Goldberg, Tobin Rower, Nathan Rower, and Lamia Ateshian) is funny, creative, and intelligent beyond words.
%
He has taught me countless invaluable lessons about being a clear and kind human being.
%
To say that his infectious positivity has changed my life is an understatement.
%
J Rogers is always, \textit{always} there when I need them, and whose incredible clarity and kindness is something I aspire to and benefit from every day.
%
Neeraj Tata, my hobbit-in-crime, has an incredible (and intentionally trained) ability of using sharp awareness and tolerance of all things to see what is really important in this life.


My academic siblings have also been important parts of my journey, and have taught me and been with me through both trials and tribulations.
%
Sean Benevedes, who I have known since my time at UCSB, carries with him an intelligence, humor, and empathy that make world a measurably better place.
%
I have also benefited from many useful conversations with Rikab Gambhir, whose funny jokes, intelligence, and unimpeachable sense of academic adventure have made my experience at MIT much more interesting.
%
I am grateful for many enthusiastic conversations about physics and life with Sokratis Trifinopoulos, and for being able to glimpse his creativity and his extremely honest positivity -- he speaks his mind even in disagreement, but somehow I always feel happier after talking with him than I had before.
%
Last but not least, Cari Cesarotti is a hilarious, high-powered, creative scientist who I love to watch playing the game of science and winning.



\vspace{10pt}
\hrule
\vspace{10pt}



The remainder of my acknowledgments are mostly chronological, and no less heartfelt.


I want to thank some of my first and most long lasting friends:
%
Liam Bradley, Daniel Kreizberg, Nate Gardner, Kyreija Lamothe, Vince Fleming, Ryan Croy.
%
Also there to support me during many formative moments were Rose Gordon, Erin Neitzel, Jason Georgis, Johannes Klein, and especially Nicholas Maguid.



More of my academic family from UCSB who have had an incredible positive influence on my life are James Ehrets (I have learned and gained so much from his openness and adventurousness, empathy, and many long conversations), Nicholas O'Dea (who perhaps belongs in my ``mentor'' section: he has shaped my entire attitude towards learning and rigor), David Newsom (he knows the many odd and memorable adventures we have shared together -- what an incredible and interesting life), Jasmin Kwak (who is an unshakeable source of support whenever I need her), Ben Siegel (whose gentle kindness I always aspire to), Dolev Bluvstein (who is a damn superhuman -- I don't know how he does it, but I feel so lucky to learn from his incredible positivity and brilliant approach to all of life), Anoop Praturu, Ryan Stuart, and Marvin Qi.
%
My b-boy crew, including Kai Goh, Eugene Ma, Brandon Louie, and Alvin Ong have given me many adventures as well, along with companionship, growth, a beautiful and positive environment for my studies, and the incredible gift of \emph{art}.
%
Yurim Lee gave me so much kindness and support through my studies at UCSB for which I am forever grateful.


At MIT, I was lucky enough to meet even more incredible and brilliant people.
%
I'm very grateful to Zhiquan Sun, who taught me among many wonderful lessons that a happy life means eating well.
%
Another close friend is Christopher McNally, who is open to anything, saved my life in the wilderness on one occasion, and saved my spirit in and out of the halls of MIT many more.
%
Some of my first friends at MIT were Atakan Firat (a tireless dynamo of intellect and creativity who never backs down from a fight), Ouail Kitouni (humble beyond belief, whose openness and kindness are only emphasized by his incredible knowledge, curiosity, and problem-solving ability), Shankar Balasubramanian (a pure genius, plain and simple, always with time to teach, share a problem, or give advice), Margarita Davydova (a kind, creative, brilliant theorist who makes science as beautiful as art), Ali Ghorashi (always with support and empathy to share), and Anjie Gao (who taught me, among other things, the wisdom that ``nothing is trivial'').
%
Though it took one more semester, I soon met Patrick Oare, Wenzer Qin, Artur Avkadiev -- my moral support group -- who have since been my people:
%
my confidants, my comrades-in-arms, my rocks.
%
Nick Kamp and Caolan John also deserve a special mention here;
%
they bring positivity and curiosity into every room, and every room is better for their presence.
%
Nathan Everett (my platonic soul mate) and Lubah Lerner (his soul mate) have also been close friends and wonderful sources of joy and support, as have Robbie Ball and Signe Wurtz.
%
I have been lucky enough to learn from and share time with Dimitra Pefkou, \r{A}smund Folkestad, Gregory Ridgway, Joshua Lin, Yitian Sun, Sarah Geller, Jamison Sloan, Robert Jones, Manu Srivastava, Kyle Lee, and Ankita Budhraja.
%
I especially want to thank Lisa Yang, whose constant drive for self-improvement is continually inspiring.


The entirety of the Engineering Quantum Systems (EQuS) ``chill team'' deserves a special mention for the joy they have brought to my life:
%
Sarah Muschinske, Miguel and Lean\'or Moreira, Jorge Marques, and Andreia Allen -- and also to Will Oliver, whose incredible positive and guiding influence has brought these wonderful, brilliant people together.
%
David Rower, Lamia Ateshian, and Dolev Bluvstein deserve more thanks here as well, and Marcin Kalinowski should be thanked three times (once for his wonderful personality, twice for his incredible science, and three times for his enormous muscles, which count as another person on their own).


Finally, I remember Jasper Pont, whose ringing laugh and free spirit still float through the world.

\vspace{12pt}
\hrule

\begin{center}
    \textit{Ashes to ashes, dust to dust.}
\end{center}
